StackTokens: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex

\begin{itemize}
\item informally explain how an adversary may try to abuse stack and return caps
\item informally explain how we prevent this using linear capabilities
\item use the tikz pictures from the PriSC presentation to explain all of this
\end{itemize}
\section{A Capability Machine with Sealing and Linear Capabilities}
\begin{itemize}
\item present our \emph{target language} and its operational semantics (excerpts)
\item mention roughly what components look like
\end{itemize}
\section{Formulating Security with a Fully Abstract Overlay Semantics}
\begin{itemize}
\item present our source language, its operational semantics (excerpts)
\item mention our assumption of reasonability
\item present the full abstraction theorem.
\end{itemize}

@@

\section{Discussion}
\begin{itemize}
\item mention that tail calls are supported through xjmp
\item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
\end{itemize}

\section{Related Work}
\begin{itemize}
\item CFI and friends
\item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
\item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
\item CheriBSD etc.
\end{itemize}
\bibliography{references}

>>>>

\begin{itemize}
\item informally explain how an adversary may try to abuse stack and return caps
\item informally explain how we prevent this using linear capabilities
\item use the tikz pictures from the PriSC presentation to explain all of this
\end{itemize}

\section{A Capability Machine with Sealing and Linear Capabilities}
\label{sec:capab-mach-with}
\begin{itemize}
\item present our \emph{target language} and its operational semantics (excerpts)
\item mention roughly what components look like
\end{itemize}

\section{Formulating Security with a Fully Abstract Overlay Semantics}
\label{sec:form-secur-with}
\begin{itemize}
\item present our source language, its operational semantics (excerpts)
\item mention our assumption of reasonability
\item present the full abstraction theorem.
\end{itemize}

@@

\section{Discussion}
\label{sec:discussion}
\begin{itemize}
\item mention that tail calls are supported through xjmp
\item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
\item Performance?  Implementability of linear caps?
\end{itemize}

\section{Related Work}

Capability machines originate with Dennis and Van Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to Levy~\cite{Levy1984capability} and Watson et al.~\cite{Watson2015Cheri} for an overview of previous work.
The capability machine formalized in \sectionname~\ref{sec:capab-mach-with} is modeled CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
The latter is a recent and relatively mature capability machine, which combines capabilities with a virtual memory approach, in the interest of backwards compatibility and gradual adoption.
In the interest of simplicity, we have omitted features of CHERI that were not needed for modelling StackTokens (e.g.\ local capabilities, virtual memory).

Plenty of other papers enforce well-bracketed control flow at a low level, but most are restricted to preventing particular types of attacks and enforce only partial correctness of control flow.
This includes particularly the line of work on \emph{control-flow integrity}~\cite{abadi_control-flow_2005}.
This technique prevents certain classes of attacks by sanitizing addresses before direct and indirect jumps based on static information about a program's control graph and a shadow stack.
Contrary to StackTokens, CFI can be implemented on commodity hardware rather than capability machines.
However, its attacker model is different and its security goals are weaker.
They assume an attacker that is not able to execute code, but can overwrite arbitrary data at any time during execution (to model buffer overflows).
In terms of security goals, the technique does not enforce local stack encapsulation.
Also, it only enforces a weak form of control flow correctness, saying that jumps stay within the static control flow graph of a program~\cite{Abadi2005Theory}.
Such a property ignores temporal properties and seems hard to use for reasoning.

More closely related to our work are papers that use separate per-component stacks, a trusted stack manager and some form of memory isolation to enforce control-flow correctness as part of a secure compilation result~\citep{patrignani_modular_2016-1,juglaret_beyond_2016-1}.
Our work differs from theirs in that we use a different low-level security primitive (a capability machine with local capabilities rather than a machine with a primitive notion of compartments) and we do not use per-component stacks or a trusted stack manager, but a single shared stack and a decentralized calling convention based on linear capabilities.
TODO: adapt the following sentence...
Also, both prove a secure compilation result from a high-level language, which clearly implies a general form of control-flow correctness, while we define a logical relation that can be used to reason about specific programs that rely on well-bracketed control flow.

\begin{itemize}
\item CFI and friends
\item CheriBSD etc.
\item our ESOP 2018 paper \citep{skorstengaard_reasoning_2017}
\item undefined behavior and full abstraction \citep{juglaret_beyond_2016} \dominique{TODO}
\end{itemize}
\bibliography{references}
